### 2021-07-07 erc-B1/parts/eatr.tex
Initial commit

% ! TeX root = ../erc-B1.tex
\subsection{Bibliometrics and summary of publication activities}

\subsection{10 research highlights}

I now present my achievements through a selection of my publications, in
reverse chronological order. My name appears in bold, names of PhD
students I supervised in italics.

\begin{asparaitem}
\item \todo{Relevant publications, with context} 
% \fullcite can be used to fully cite a publication
\end{asparaitem}

\subsection{Invited presentations}

\noindent\hskip -.5cm\begin{tabular}{>{\bfseries}rll}
  \toprule
  \multicolumn{1}{c}{\bfseries Date}&{\bfseries Location}&{\bfseries
Context or inviting institution}\\
  \midrule
  ~\\
  \bottomrule
\end{tabular}

\subsection{Awards}