### 2021-11-29 paper-2021-woa-ac-ml.tex
final Musolesi pass

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{cite}

\usepackage{tcolorbox}
\usepackage[colorinlistoftodos,prependcaption]{todonotes}
\usepackage{xcolor}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{minted}
\usepackage{subcaption}
\usepackage[draft]{hyperref}
\usepackage[inline]{enumitem}
\usepackage[T1]{fontenc}
\usepackage{cleveref}
\usepackage{systeme}

\setminted[scala]{baselinestretch=1,bgcolor=white,frame=single}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
  T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\newcommand{\unibo}{\textsc{Alma Mater Studiorum} -- \textit{Università di Bologna}}
\newcommand{\disi}{\textit{Computer Science and Engineering - DISI}}
\newcommand{\scafi}{{ScaFi}}
\newcommand{\scalaloci}{{ScalaLoci}}
\newcommand{\scafiloci}{{ScaFiLoci}}
\newcommand{\scalainline}[1]{\mintinline[fontsize=\small]{scala}{#1}}
%%%%% LEGENDA FOR NOTES %%%%%
%% RED: constraints
%% YELLOW: suggestions
%% CYAN: todos
%% GREEN: outline
\newcommand{\constraints}[1]{\todo[inline, color=red]{#1}}
\newcommand{\suggestions}[1]{\todo[inline, color=yellow]{#1}}
\newcommand{\todos}[1]{\todo[inline, color=cyan]{\textbf{TODO}: #1}}
\newcommand{\outline}[1]{\todo[inline, color=green]{#1}}
%% Utility commands
\newcommand{\round}{\texttt{round}}
\newcommand{\export}{\texttt{export}}
\sloppy
\begin{document}
\title{Towards a Reinforcement Learning approach for Aggregate Computing} %% or Towards a Machine Learning approach for Aggregate Computing
\author{
\IEEEauthorblockN{
Gianluca Aguzzi\IEEEauthorrefmark{1},
Roberto Casadei\IEEEauthorrefmark{1},
Mirco Musolesi\IEEEauthorrefmark{1}\IEEEauthorrefmark{2},
Mirko Viroli\IEEEauthorrefmark{1}
}
\IEEEauthorblockA{\IEEEauthorrefmark{1}
\textsc{Alma Mater Studiorum}---Universit\`a di Bologna, Cesena, Italy}
\IEEEauthorblockA{\IEEEauthorrefmark{2}
University College London, London, United Kingdom\\
Email: \{gianluca.aguzzi, roby.casadei, mirco.musolesi, mirko.viroli\}@unibo.it
}}

%% Minted helpers
\definecolor{mblue}{rgb}{0.27,0.33,0.53}
%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
%\begin{keywords}
%  Collective-Adaptive System \sep
%  Aggregate Computing \sep
%  Deep Learning \sep
%  Reinforcement Learning \sep
%\end{keywords}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.

\maketitle
\begin{abstract}
Recent trends like pervasive, autonomic, and swarm computing
 promote 
 a vision of large-scale collections of relatively simple agents
 that act and coordinate 
 with no central orchestrator
 to support distributed applications.
%
Such Collective Adaptive Systems (CAS)
 are hard to engineer,
 because mechanisms for developing global behaviour
 out of local behaviour and interaction must be devised,
 often leveraging natural inspiration.
%
In particular, in the field-based Aggregate Computing approach,
 collective adaptive behaviour is specified in terms of computational fields,
 namely maps from CAS entities to computational values evolving over time, and neighbour-based interaction.
%
However, devising efficient aggregate computing algorithms
 for different scenarios
 is not an easy task; in particular,
 a problem is determining the best local decisions based on the current context
 for fostering collective outcomes.
%
Therefore, in this work we investigate
 the possibility of using machine learning techniques
 to automatically devise local actions
 to improve existing aggregate algorithms.
%
In particular, we propose a Reinforcement Learning solution that can be seen as the first step towards the integration between Aggregate Computing and Machine Learning. 
% The goal is to improve adaptivity through low-level specifications but maintaining high-level program behaviours. 
Accordingly, we outline challenges and opportunities of this integration, and present preliminary results based on a case study where Aggregate Computing is successfully combined with Hysteretic Q-Learning.
%

\end{abstract}
\begin{IEEEkeywords}
  Collective-Adaptive System,
  Aggregate Computing,
  Deep Learning,
  Reinforcement Learning
\end{IEEEkeywords}

\section{Introduction}
Recent trends like pervasive~\cite{DBLP:journals/wc/Satyanarayanan01} and autonomic~\cite{DBLP:journals/computer/KephartC03} computing 
 promote a vision of artificially designed Collective Adaptive Systems (CASs)~\cite{DBLP:conf/huc/Ferscha15},
 namely collections of agents coordinating and adapting %as wholes 
 to pursue joint goals in changing environments.
% due to the increasing interconnected computational devices placed in ever-changing environments.
 Examples of CASs include robot swarms, smart cities, and digitally-augmented crowds.
%
%CASs are characterised by 
%\begin{itemize}
%\item a global order that emerges from local node interactions, 
%\item  decentralised control, and 
%\item adaptive behaviours towards environmental changes to pursuit a collective goal.
%\end{itemize}
%
CAS are notoriously hard to design and control 
 %due to as part of complex systems
 as globally desired behaviour
 has to emerge from local behaviours and interactions. 
%
For this reason, over the years, various approaches have addressed the problem of dealing with this complexity---for instance, by taking inspiration from natural systems like insect and animal societies (e.g., ant colonies, bee hives, flocks of birds etc.). 
%
%In this case, a sort of hive mind arises 
% %-- also referred to as swarm intelligence --  
% born from local interactions between agents. 
%
Specifically, such approaches tend to exploit 
 \emph{self-organisation}, 
 i.e., processes by which global order emerges 
 from continuous interactions among simple entities.

Initially, researchers had tried to build systems that show these properties in computing systems by mimickng natural collective behaviours ``from the bottom up''~\cite{DBLP:journals/connection/Webb02}, 
 namely trying to find the right single-node-behaviour (local rules)
 to achieve the intended global behaviour. %, e.g. flocking, foraging, etc.
%
However, such local rules cannot easily be defined, since emergence, and its effects, are (almost by definition) hard to be controlled or predicted in complex systems.
%
A different approach could attempt to express collective behaviours in a more top-down fashion, in terms of the desired emergents.
%, abstracting over non-functional aspects such as the network topology, environment uncertainties and system size.
In this direction, Aggregate Computing (AC)~\cite{DBLP:journals/computer/BealPV15} is proposed:
 it is a paradigm by which
 developers can express collective self-organising behaviours from a global perspective. % through a functional program specification.
%
Indeed, an aggregate program 
 (i.e. a program written with AC) 
 consists of the manipulation of \textit{computational fields}, i.e., distributed
 data structures where each node is associated with the result of a computation performed by it,
 and neighbourhood observation.
%
%These operations are defined in \textit{field-calculus}~\cite{DBLP:conf/coordination/AudritoBDV18} --- the root of Aggregate Computing. 
% This calculus describes a set of minimal constructs to express any spatiotemporal computation.

AC captures common abstractions for collective behaviour specification in so-called ``building-blocks''. 
 %These abstractions are encapsulated in so-called ``building blocks''.
 %through which various applications have been built (e.g., for swarm robotics and crowds engineering~\cite{DBLP:journals/eaai/CasadeiVAPD21}).
However, building block synthesis concerns different elements not only related to the behaviour itself
 but also to the ``ensemble" specifications. 
%
For instance, it is quite difficult to craft building blocks
 that easily adapt to the non stationarity due to node movements.
%
Ideally, we would like to continue developing our applications using these building blocks 
 -- so the high-level specification remains the same -- 
 but using some under-the-hood mechanisms that adapt the behaviour according to different environmental changes. 
%
For this reason, we claim that Machine Learning can help us in handling this "unseen" situation, leaving only the burden of defining the right collective specifications to developers.
%This lead to an ``hybrid" approach where a part of behaviour is still expressed using common 
% AC abstractions and another part is distilled through learning~\cite{research}.
In particular, in this paper,
 we investigate how AC can be extended with Reinforcement Learning (RL) through a case study,
 where we outline the performance improvement of the ``hop count" program ---
 i.e. an aggregate program that produces a field that stores the distance from a source zone. 
 Furthermore, we show that the learnt policy scales up to different system sizes.

The paper is organised as follows. 
%
In \Cref{background} we describe
 the AC framework, %showing the context in which is typically applied and the overall idea of how it works, and then the motivation of injecting Reinforcement Learning in this paradigm,
 and motivate our proposal to inject RL into this paradigm. 
 %in \Cref{aggregate-and-machine-learning} we detail the relevant properties of Aggregate Computing in terms of machine learning problems, 
 In \Cref{aggregate-and-rl}, we describe the details of our integration of AC with RL.
Then, in  \Cref{evaluation} we evaluate the approach on a paradigmatic AC case study.
%
Finally, in \Cref{conclusion} we discuss the results and the path towards deeper AC-RL integration.

 \section{Background and Motivation}\label{background}

%Computer Science is an incredibly evolving research area. 
% In a few years, the environment had been filled with intercommunicating computational devices, such as smartphones, smart things, and personal computers.

Collective Adaptive Systems are characterized by a very high number of nodes without a central authority that coordinate nodes. 
 The overall behaviour is a consequence of local interactions.%The systems expose self-* properties.

Engineering this kind of system is a thriving area, and so various approaches have been proposed.
 First attempts leverage \textit{local-to-global} techniques, where the collective behaviour emerges from local behaviour description, 
 usually inspired by nature~\cite{DBLP:journals/swarm/BrambillaFBD13}. 
%
Instead, a natural way to describe this kind of problems should be through a \textit{collective} view of the system.
 This is the idea used by the so-called \textit{global-to-local} techniques~\cite{DBLP:journals/jlap/ViroliBDACP19,DBLP:journals/scp/AlrahmanNL20, DBLP:conf/cbse/BuresGHKKP13}, 
 where the behaviours are designed \textit{top-down}, and then the local one is generated automatically from this ``system-wide'' specification. 
% 
Also ``meet-in-the-middle''~\cite{DBLP:journals/computer/PinciroliB16} approaches exist, where
 a part of the system is still expressed bottom-up (device-centric) whereas a part is expressed top-down (swarm-centric)

In this work, we focus on Aggregate Computing~\cite{DBLP:journals/computer/BealPV15}, a global-to-local approach where
 the self-organising collective behaviour is expressed in terms of manipulation of \textit{computational fields}.
The latter could be seen as the application of the mathematical concept of field: it consists of a 
 map from any devices to their computational value (i.e., the result of a computation).
The computational field manipulation is ruled by \textit{field calculus}, which define the constructs needed 
 to express spatio-temporal computations~\cite{DBLP:journals/jlap/ViroliBDACP19}.
%
The novelty of AC resides in its composable (functional-like) system-size independent program definition: 
 using basic constructs we identify the main patterns (\textit{building-blocks}) that could be reusable at application level.
By opportunistically crafting these building-blocks, it is possible to verify prominent properties such as eventual consistency %~\cite{DBLP:journals/taas/BealVPD17} 
 (i.e. the ensemble behaviour is independent of the underlying network details)
 and self-stabilisation %~\cite{DBLP:journals/tomacs/ViroliABDP18} 
 (i.e. the ability of a system to recover from arbitrary changes)~\cite{DBLP:journals/jlap/ViroliBDACP19}.

AC is studied in various simulated scenarios ranging from smart cities to crowd management~\cite{DBLP:journals/jlap/ViroliBDACP19}, 
 thanks to Alchemist~\cite{DBLP:journals/jos/PianiniMV13} (a meta-simulator for pervasive systems) and programming languages and tools that support 
 this paradigm~\cite{DBLP:journals/jlap/ViroliBDACP19}, such as the ScaFi aggregate programming language~\cite{DBLP:conf/isola/CasadeiVAD20}. %, FCCP~\cite{DBLP:conf/acsos/Audrito20} and Protelis~\cite{PianiniSAC2015}.

In the next sections, we summarise the field calculus language and computation model, which are essential to understand the case study of \Cref{evaluation}.
Then, we discuss the motivation that brought us to integrate Reinforcement Learning into the AC framework.

\subsection{Field Calculus operators}
The main constructs that capture the essential aspects for programming self-organising systems with field calculus are:
\begin{itemize}
  \item \textit{Stateful field evolution} expression \mintinline[escapeinside=||]{scala}{|{\color{teal}rep}|(|$e_1$|) {|($x$) => $e_2$|}} describes a field evolving in time. 
  $e_1$ is the initial field value and the function $(x) => e_2$ define how the field changes at each execution.
  \item \textit{Neighbour interaction} expression \mintinline[escapeinside=||]{scala}{|\color{teal}nbr|{|$e$|}} a view
  of the field values in the surroundings of each device where neighbours are
  mapped to their evaluations of $e$. A series of \mintinline{scala}{*hood} operators leverage this map to produce relevant result. 
  For instance, \mintinline{scala}{minHood} returns the minimum value 
  within the neighbourhood.
  \item \textit{Domain partitioning} expression \mintinline[escapeinside=||]{scala}{|\color{teal}if|(|$e_0$|){|$e_1$|}{|$e_2$|}} splits the computational
  field into two non-communicating zones hosting isolated subcomputations:
  $e_1$ where $e_0$ is true, and $e_2$ where $e_0$ is false.
\end{itemize}
%
A full coverage of field calculus is beyond the scope of this paper. A more comprehensive presentation is given in~\cite{DBLP:journals/jlap/ViroliBDACP19}.

\subsection{Computational Model}
%
For an aggregate program to yield collective adaptive behaviour, 
 continuous computation and communication are needed.
%
An individual atomic step 
 involving context acquisition,
 computation of the program,
 and neighbourhood-based communication 
 is called a \round{}.
%
In detail, a \round{} run at some device consists of:
\begin{enumerate}
  \item \emph{context acquisition}: the device collects information from the sensors, 
  the messages from neighbours
  (including the device itself---to model state);
  \item \emph{program evaluation}: the aggregate program is evaluated against the acquired context,
  yielding an \export{}, namely the message to be sent to neighbours;
  \item \emph{export sharing}: the \export{} is sent to the neighbours;
  \item \emph{actuations}: the \export{} also includes information that can be used to drive actuators in the local node.
\end{enumerate}

Rounds execution is completely asynchronous, so there is no global clock or barrier to coordinate the aggregate. 
%
 This makes it possible to scale easily with the number of nodes. 
%
The combination of the logic of the aggregate program
 and 
 such a collective, proactive, and periodical execution
 can promote the pursuit of collective adaptive behaviour.

\subsection{Motivation}

Consider the problem of associating
 to each node of the network
 a value
 representing the estimation of the minimum distance
 from that node to the closest \emph{source} node (see \Cref{fig:hop-count}).
%
This is indeed a computational field 
 and it is known as a \emph{gradient}~\cite{DBLP:conf/saso/AudritoCDV17}
 and there exist aggregate algorithms
 for computing it in such a way that
 the field automatically adapts 
 to changes induced by mobility
  and addition/removal of sources.
%
This example is significant as the gradient is a key building block 
 for more complex self-organising behaviour~\cite{DBLP:journals/jlap/ViroliBDACP19}.

%Currently, aggregate computing has a problem in the ability to generalise collective behaviours that are capable of working well in different scenarios. 
%%
% For example, a well-tested aggregate application in an environment with limited dynamics could lead to a non-stable behaviour under continuous perturbations.
%% 
% Typically, to alleviate this problem, designers create ad-hoc building blocks that cope with certain environmental aspects (such as the ULT block~\cite{DBLP:conf/saso/AudritoCDV17}). 
% But they are still context-specific solutions.
However, in practice, 
 specific gradient algorithms
 may be developed
 to achieve different trade-offs 
 in terms of reactivity to change and stability~\cite{DBLP:conf/saso/AudritoCDV17},
 which make them more or less suitable depending on 
 the environmental and network dynamics at hand.
%
The problem is that manually crafting such algorithms
 is not an easy task, for various heuristics and assumptions typically have to be considered.
%
To address this problem,
 we propose to use machine learning techniques
 in order to learn ``good'' local decision-making.
% 
That is, the idea is to leave the low-level aspects of AC to Machine Learning algorithms, 
 training the collective in different environments and thus building general behaviours that are capable of adapting to many circumstances.

In this case, when Reinforcement Learning is applied to several agents, we speak of Multi-Agent Reinforcement Learning (MARL)~\cite{DBLP:journals/aamas/Hernandez-LealK19}.
 AC has a typical cooperative and homogeneous learning setting: 
 each agent collaborates to achieve a common goal guided by the same reward function.
%
In the cooperative and homogeneous systems, several works have tried to apply the algorithms used in the single-agent context to the multi-agent one --- 
 the so-called Independent MARL~\cite{DBLP:journals/tsmc/BusoniuBS08}. 
%
However, several issues emerge, such as non-stationarity, multi-agent credit assignment, scalability and partial observability of the system. 
 Therefore, several heuristics have tried to solve these problems.
%
Among many, Hysteric Q-Learning~\cite{hysteretic-q} uses an optimistic approach in which more weight is given to positive actions than negative ones.
%
Another way to handle multi-agent learning is to use the so-called \textit{Centralised Training and Decentralised Execution} (CTDE) algorithms. 
 In this case, a central agent collects the experiences of the various agents and then distils a policy that each agent can execute locally in a typical actor-critic setting. 
%
During the learning phase, the problems of partial observability and non-stationarity are reduced, since the centralised learner has a global view. 
 However, the problem of scalability is typically not addressed.

In the next section, we exploit Hysteretic Q-Learning (since it empirically showed good performance~\cite{DBLP:journals/tsg/HadidiJ13}) with AC, 
 as the first step towards \textit{``hybrid"} techniques for programming aggregate systems~\cite{research}.
 In this way, an application part is described in terms of aggregate computations and the low-level part is left to learning.
% \section{Aggregate Computing and Machine Learning}\label{aggregate-and-machine-learning}

% In this section, we analyse the characteristics of the AC paradigm,
%  map these to relevant machine learning contributions in the literature,
%  and point out different learning perspectives about the integration of AC with Machine Learning

% \paragraph{Multi-and many-agent system}
% %
% An aggregate system is a multi-agent system
%  and, often, a \emph{many}-agent system
%  where a large number (hundreds or more)
%  of autonomous entities are programmed to achieve 
%  some collective behaviour through \emph{repeated} 
%  sensing, computation, communication, and actuation steps.
% %
% Due to the high stochasticity of the environment,
%  it is almost impossible to know and
%  program the optimal behaviour for all agents in advance.
%  This uncertainty results in the need of creating intelligent agents
%  so that they can learn the optimal behaviour and adapt to environmental changes.
% %
% In recent decades there has been an emerging trend in the use of RL 
%  in multi-agent settings -- called Multi-Agent Reinforcement Learning (MARL) -- as a powerful, robust and adaptive learning paradigm.
%  Progress has been considerable and a wide range of algorithms are now available.

% MARL is a conjunction of Game theory and RL, 
%  and there are several (even orthogonal) viewpoints on which researchers have been focused.

% First attempts from RL viewpoint, 
%  goes towards a so-called \textit{independent learning}~\cite{DBLP:journals/tsmc/BusoniuBS08} approach, 
%  where each agent learn locally against the whole environment~\cite{DBLP:conf/icml/Tan93}.

% Examples of these are Hysteretic Q-Learning (used to reduce non-stationarity)~\cite{hysteretic-q}, 
%  Lenient Learning~\cite{DBLP:journals/jmlr/WeiL16}, and Frequency Maximum Q-value~\cite{DBLP:conf/atal/KaisersT10}.
% %
% The pro of these approaches is that their complexity does not scale up with the number of agents, 
%  but, unfortunately, the learning process is extremely non-stationary and unstable.
% % 
% Furthermore, they are heuristic and do not exist any convergence proof (even if, in practice, they often reach a good policy).

% Other efforts have focussed on achieving equilibria such as Nash-equilibrium of Pareto optimality.
%  The first works towards this direction are Nash-Q-Learning~\cite{nash-q}, Friend-or-Foe Q-learning~\cite{DBLP:conf/icml/Littman01}, Minimax-Q~\cite{DBLP:conf/icml/Littman94}.
% The problem is that it is used with few agents and scale hardly with the application complexity.
% %
% Finally, current emerging trends tend to leverage a so-called centralised training and decentralised execution (CTDE) approach, by which 
%  agents should leverage the system-wide knowledge at training time but at runtime, they act independently~\cite{DBLP:journals/aamas/Hernandez-LealK19}.
%  Most promising approaches in this direction are MAPPO~\cite{DBLP:journals/corr/abs-2103-01955}, COMA~\cite{DBLP:journals/corr/FoersterFANW17} and MADDPG~\cite{DBLP:conf/nips/LoweWTHAM17}.

% The main problem with most of the solutions available in the literature is that they consider a small number of agents (or at least test them on small games).
%  In AC instead, we need to find a technique that does not depend on the system node count.

% \paragraph{Neighbour-based or indirect (environment-mediated) interaction}
% %
% In aggregate systems, a device can directly interact only with its neighbours.
% %
% Data flows may be implemented
%  to support indirect communication across multiple hops;
%  however, information from devices far away tends to be obsolete.
% %
% The environment can also be used, via stigmergy~\cite{DBLP:journals/scp/NicolaSI20},
%  for indirect communication;
%  however, a device can typically only access 
%   to a very local portion of the environment.
% %
% In other words, the system state is only \emph{partially observable}.
% In practice, CTDE technique aims to solve these problems, using a \emph{global} view at training time but
%  a \emph{partial view} at execution time.

% Furthermore, when the system node size is very large, novel works of Mean-Field Reinforcement Learning~\cite{DBLP:conf/ijcnn/ZhouZX21} are able to
%  handle learning efficiently, reasoning about agent and agent neighbourhoods.
% %
% \paragraph{Learning goal: collective behaviour}
% %
% The devices of an aggregate 
%  must cooperatively learn the ``aggregate program'',
%  namely the collective behaviour 
%  that achieves a particular \emph{global goal}
%  in a decentralised, resilient way.
% %

% % \paragraph{Self-organisation, swarm robotics} it is condensed it other perspective, so I think it is better to avoid this paragraph.
% %
% % \suggestions{You can talk about similar problem structure (i.e. the collective follow the same local behaviour)}

% \paragraph{Emergent behaviour}
% %
% The collective behaviour of an aggregate system
%  \emph{emerges} 
%  from a dense network of computations and interactions
%  in an evolving environment.
% %

% Typically, in CAS, single-agent RL brings to emergent behaviour and is used in various applications~\cite{DBLP:conf/icse/DAngeloGGGNPT19}.

% \paragraph{Delayed, global rewards}
% %
% In the light of the above properties,
%  if we consider a learning approach based on rewards,
%  we must observe that such rewards would be
%  delayed and mostly global---i.e.,
%  it may not be easy to solve the credit assignment problem
%  to distinguish good from bad individual contributions
%  to the global outcome.

% A way to handle this kind of rewards in multi-agent settings is \textit{difference rewards}~\cite{DBLP:conf/atal/DevlinYKT14, DBLP:conf/aaai/FoersterFANW18}. It is used to capture agents contributions to the system's global performance.
% \paragraph{Transient phase}
% %
% When an aggregate behaviour is carried out,
%  we often focus on the final stable outcomes after the transitory period.
% %
% However, intermediate results are also important,
%  and invariants should hold throughout the entire computation.

% \paragraph{Dynamic topology}
% %
% If we admit \emph{mobility} and \emph{failure} (rarely considered in related works, where node population is fixed),
%  then we must also consider
%  the possibility of changes in neighbourhoods
%  and hence the need of dealing with dynamic topologies.

% \subsection{Integrating Learning Mechanisms into Aggregate Computing}
% We now outline the different perspectives about AC with learning, 
%  to clarify the current and the future research path.
% %
% \paragraph{Application} Learning can be useful at different levels in our framework.
%  First of all, learning can be applied to the building blocks definition.
% % 
% Indeed, most of our results and applications are based on them.
% %
% The building block improvement may consist either of learning a correction factor or by entirely learning the behaviour.

% On the other side, learning can be useful also for framework related stuff,
%  mainly aiming at "non-functional" aspects. 
%  Trust mechanisms, evaluation frequency, energy consumption manager, package storage are all the things that could take advantage of using Machine Learning algorithms.

% \paragraph{Learning problem} 
% The learning problem in multi-agent settings can be posed as \textit{Independent} and \textit{Team} learning. 
% In the foster, each agent perform the learning process and while in the latter,
%  the learning process is performed only by one agent and then shared with the collective (in a typical CDTE setting).
% % 
% Team learning is usually used in the ``swarm'' system, due to the homogeneous behaviour 
%  --- i.e. each agent performs the same program. 
%  Here, a common reward signal exists and evaluate the overall collective behaviour.
% % 
% In AC, each agent follows the \textit{same} aggregate program shared within the entire system, so Team learning can 
%  be applied seamlessly.

% On the other hand, Independent Learning is usually applied in Collective
%  (Self-)Adaptive System \cite{csas-and-marl}.

% In particular, Independent Reinforcement Learning reach good results in this kind of system, 
%  even if no theorem exists as in the single-agent context.
% %
% With Independent learning is still possible to reach a global collective behaviour convergence. 
%  Indeed using a common reward signal, the collective may reach the same behaviour through emergence \cite{iima2008swarm, nguyen2018swarm}.

% \paragraph{Learning technique}
% Aside from what kind of problem we set up, we can then use different Machine Learning techniques,
%  such as RL or Supervised Learning (SL).
% %
% SL is rare in CSAS -- or in the ``swarm" like system in general.
% %
% Novel works~\cite{DBLP:conf/corl/TolstayaGPP0R19} leverage Graph Neural network to learn how agents should communicate, 
%  supposing to know the right behaviour (i.e. produced via global vision simulation or by emulation of natural phenomena).

% In AC, S) can be applied when we want to build up new building blocks from zero. 
% % 
% Indeed, we eventually know the right state of the system, but we do not know how the system can reach it.

% RL is predominantly used in CAS, 
%  due to its flexibility and thanks to the plethora of approaches used in this scenario~\cite{csas-and-marl}.

% In our case, RL can be used both at the building block level and the framework level.
% %
% For the foster, the simplest scenario can be that agents learn a correction factor to improve a long-term reward signal.

% \paragraph{Learning time}
% Finally, learning can be performed \textit{offline}, \textit{online} or \textit{offline an then online} (\textit{mixed time}).
% % 
% The foster is the common scenario in multi-agent settings.
%  Here indeed, we can leverage simulation to know the right computational field and then use one of the novels CTDE techniques. 
% Mixed-time learning can be a mid-level complexity direction instead.
%  Here, simulation can always be exploited to reach a good behaviour, but at runtime, 
%  the behaviour can be adapted based on the new environmental situations. 
% Finally, we can imagine that learning can be performed online. 
%  It is a common situation where we do not know the environment and also simulation can be very hard to accomplish. 
%  In mixed and online cases, RL is the most suitable model, because agents \textit{learn by doing}.

\section{Aggregate Computing with Reinforcement}\label{aggregate-and-rl}
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{img/algorithm-learning.pdf}
  \caption{This figure shows the main steps of RL in AC.
  Each agent starts with his copy of a Q-table. 
  Then, during the simulation, they refine them Q-tables changing Q values according to their experience (different colours mean different values). 
  In the end, a new $Q_{max}$ table will be created aggregating each agent table. 
  Finally, a new episode starts using this $Q_{max}$ as a starting Q-table.
  }
  \label{fig:aggregate-q-learning}
\end{figure}
%As first experiments, we leverage RL to improve building blocks. 
% 
In this section, we briefly describe how RL can be injected into the AC computation loop,
 and what RL we used with AC.

From a local point of view, AC program execution can be seen as:
$$
f : \textit{Context} \rightarrow \textit{Output}
$$
Where $\textit{Context}$ contains:
\begin{itemize}
  \item $\sigma(name) \rightarrow \textit{data}$: a function from sensor name to the perceived value;
  \item $\phi(\textit{ID}) \rightarrow \textit{Export}$: a function from id to neighbourhoods export data.
\end{itemize}
The computational field $\theta$ contains all output produced by the system.

As far as RL settings are concerned, we need to define the state function and the reward function.
%
The execution needs to be distributed, so the state will be a function from the local node context:
$$
K := \Omega(\textit{Context})
$$

Instead, reward function might be dependent on the learning technique, time, and problem. 
 Indeed, if we perform learning offline, it could be defined from the overall \textit{simulated} computational field:
$$
\textit{R}_{global} := \psi({\theta})
$$
If we use an online mode, it needs to be defined as a function to the local node context:
$$
\textit{R}_{local} := \psi(\textit{Context})
$$
Finally, actions influence the current aggregate program output, changing the program evaluation as:
$$
f_{rl} : \textit{Context} \rightarrow \lambda(\textit{Output}, \textit{Action})
$$
As remarked before, the policy $\pi$ needs to be distributed, so that each agent can choose the action that will be performed:
$$
\textit{Action} \leftarrow \pi(K)
$$

As a first step towards a complete integration between AC and RL, we have decided to adopt an Independent RL learning technique, 
 by which each agent try to improve their own current local policy (shown in \Cref{fig:aggregate-q-learning}). 
% 
We made this choice inspired by other work in CSAS~\cite{csas-and-marl} and swarm systems~\cite{nguyen2018swarm}.

Our approach is partially inspired by Swarm Q-Learning~\cite{nguyen2018swarm} and leverage Hysteretic Q-Learning~\cite{hysteretic-q} to
 handle the non-stationarity in the environment.
%
In particular, each agent updates a local Q-table ($Q_i(s, a)$, where $i$ is the agent id) following the standard hysteretic update:
$$
\delta \leftarrow r + \gamma * max_a Q_i(s', a) - Q_i(s_t, a_t)
$$
$$
Q(s_t, a_t) =  \begin{cases} 
  Q(s_t, a_t) + \alpha * \delta & \mbox{if } delta >= 0 \\ 
  Q(s_t, a_t) + \beta * \delta & \mbox{otherwise }
\end{cases}
$$

At the end of each episode, each agent collects its local experience, which can vary in different zones. 
 So, we combine the $N$ (the number of nodes) Q-table selecting the maximum value for each state-action pair since we assume an optimistic homogeneous behaviour:
$$
Q_{max}(s, a) = max_i(Q_i(s, a))
$$
Then, in the next episode, each agent starts with the $Q_{max}$ and try to refine it.
\section{Evaluation}\label{evaluation}
\begin{figure*}[h]
  \centering
  \begin{subfigure}[b]{0.32\textwidth}
      \centering
      \includegraphics[width=\textwidth]{img/hop-count-1.png}
      \caption{Stable situation: Two source nodes exists}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.32\textwidth}
      \centering
      \includegraphics[width=\textwidth]{img/hop-count-2.png}
      \caption{Unstable situation: One source disappears}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.32\textwidth}
      \centering
      \includegraphics[width=\textwidth]{img/hop-count-3.png}
      \caption{Stable state reached}
  \end{subfigure}
  \caption{Graphical result of hop count evaluation. The node with a red cross is the target zone. The label 
  associated with each node is the result of the aggregate program evaluated in that zone.}
  \label{fig:hop-count}
\end{figure*}
In this section, we outline the results of the application of RL for one simple
 aggregate program: the hop count. In this case, the computational field that will eventually be produced consists of the distance in terms of the minimum number of nodes needed to reach source nodes (\Cref{fig:hop-count}).

A naive implementation of this kind of problem can be described in AC as:
\begin{minted}[fontsize=\scriptsize,mathescape]{scala}
def hopCount(): Int = 
  // update state (initially, $\infty$; now, hopCounts) by ...
  rep(Double.PositiveInfinity) { hopCounts => 
    // source returns 0,
    // non-sources minimise over neighbourdhood and add 1
    mux(source) { 0 }{ minHoodPlus(nbr{hopCounts}) + 1 }
  }
\end{minted}
However, this solution suffers from the slow-rising problem~\cite{DBLP:conf/saso/AudritoCDV17}. 
 It consists of a slow convergence time when an equilibrium situation is broken.
So, our first intuition is to accelerate convergence through learning.
 In particular, the previous solution can be rewritten as:

 \begin{minted}[fontsize=\scriptsize]{scala}
def hopCountRl(k: Int, w: Int): Int = 
  rep(Double.PositiveInfinity) { hopCounts => 
      val currentState = 
        state(hopCounts, k, w)
      val action = localPolicy(state)
      hopCout() + action
  }
\end{minted}

Where the actions are discrete values (0, 1, 2 in our experiments) that correct the local result following a policy refined through learning.
%
The state instead, is built upon previous output and of the neighbourhoods output.
 In particular, in this case, it is built from a temporal window in order to encode the hop count velocity:
\begin{minted}[fontsize=\scriptsize]{scala}
def speed(hopCount: Int, k: Int) =
{
  val min = minHood(nbr{ hopCounts })
  val window = recentValues(min, k)
  min - window.head 
}
\end{minted}
\mintinline{scala}{recentValues(data, k)} is an aggregate operator that returns a list containing the last \textbf{\texttt{k}} data perceived by the agent.
Then, due to the partial observability of the environment, we maintain a 
 list of velocities, to understand the hop count direction:

 \begin{minted}[fontsize=\scriptsize]{scala}
def state(hopCount: Int, k: Int, w: Int) =
{
  val currentSpeed = speed(hopCount, k)
  recentValues(currentSpeed, w)
}
\end{minted}

Where \textbf{\texttt{k}} and \textbf{\texttt{w}} are two parameters of this algorithm.
%
The policy is learnt locally, during simulation, using our version of Q-Learning algorithm. 
 The reward function consists in: 
\[
\systeme*{0 : & \mbox{\textit{hop count output is right}}, -1 : & \mbox{\textit{otherwise}}}
\]
The idea is to maintain the nodes in a stable situation or to
 speed up the time needed to reach a stable one.
%
\Cref{fig:aggregate-learning-loop} shows the learning loop in details.
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{img/aggregate-program-evaluation.pdf}
  \caption{Each agent, starting from their context, 
  periodically performs the standard hop count. 
  Using the output produced and the history of state and output, 
  each agent computes its current state. 
  Then, an agent evaluates the action to speed up the hop count convergence following a policy. 
  This new refined output is then used to evaluate the reward and to improve the Q-table (dotted line path)}
  \label{fig:aggregate-learning-loop}
\end{figure}

\subsection{Simulation Setup}
We verify our learning algorithm through simulation\footnote{repository at \url{https://github.com/cric96/alpaca-2022-ac-rl}} leveraging Alchemist and ScaFi.
 To check if the aggregate eventually speeds up the hop count convergence time, we create a situation where the rising problem emerges (\Cref{fig:hop-count}):
 firstly, two source nodes exist in the network, ($A$ and $B$).
 At the time $t$, $B$ disappears. So the nodes near $B$ start a slow process
 to eventually reach the distance towards $A$.
We use 80 ($N$) nodes displaced into an irregular grid with a width of 100 meters and a height of 20 meters. 
 Each node has a neighbourhood range equal to 8 meters --- on average, each node has 7 neighbours. 
 The time $t$ is fixed to 20.
 Each simulation episode lasts 60 time-unit. Finally, we set $k = 3$ and $w = 2$.
% 
We train the aggregate for 500 episodes.
 The Q-learning parameters are:
$$ 
\alpha = 0.5,
\beta = 0.05,
\gamma = 0.9
$$
Moreover, we use a $\epsilon$ decay policy for the $\epsilon$-greedy policy as:
\begin{equation*}
  \begin{array}{l}
    \epsilon_t = \epsilon_{start} * e^{-t/\theta}\\ 
    \epsilon_{start }= 0.99, \theta=4000
  \end{array}
\end{equation*}
In doing so, the agent will explore the state-action space as much as possible.

\Cref{fig:simulation}-a shows the error for both naive implementation and RL-based implementation as time changes.
The error is evaluated as:
\begin{equation*}
  \begin{array}{l}
    error(i, t) \leftarrow  | \textit{rightValue}(i,t) - \textit{output}(i,t) | \\
    total_{error} = \frac{1}{T} * \sum_{i = 0}^T \frac{1}{N} * \sum_{j = 0}^N = error(i, j)
  \end{array}
\end{equation*}
%
where \textbf{i} is the node identifier and \textbf{t} is the time when the error is evaluated.

Then, we test our learnt algorithm in different environments (\Cref{fig:simulation} b, c, d). To do that, we share the learnt Q-table to the 
 entire aggregate and we test the program changing the system size (80, 360 and 760 nodes).
%
For each population size, we run 50 episodes and then we average the error at each time step. 
The aggregate speed-up convergence is observed for each scenario taken into consideration.

\begin{figure*}
  \begin{subfigure}[b]{\textwidth}
    \centering
    \includegraphics[width=\textwidth]{img/mean-error}
    \caption{The \textit{orange} line 
    describes the RL solution error. The \textit{blue} line shows
    the naive solution error.}
    \hspace{5mm}
    \label{fig:simulation-a}
  \end{subfigure}

  \centering
  \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{img/80}
      \caption{}
      \label{fig:simulation-b}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{img/360}
      \caption{}
      \label{fig:simulation-c}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{img/760}
      \caption{}
      \label{fig:simulation-d}
  \end{subfigure}
  \caption{Analysis of the learning process. \Cref{fig:simulation-a} shows the $total_{error}$ evolution during the learning process. \Cref{fig:simulation-b}, \Cref{fig:simulation-c}, and \Cref{fig:simulation-d} show the average error in three different simulations with 80, 360 and 760 nodes. 
  The black line is the error of the RL-based solution, 
  the green line is the error produced by the standard hop count implementation.}
  \label{fig:simulation}
\end{figure*}
%\todos{add blinking scenarios}
%\todos{add movement scenarios}

\section{Discussion and Concluding Remarks}\label{conclusion}
In this article, we have discussed the combination of AC with Machine Learning techniques in order to improve adaptivity
 and effectiveness of the AC framework.
In particular, we explore the combination of AC with Hysteretic Q-Learning.
The application of RL to this kind of setting is a largely unexplored %-- even if some works exists~\cite{DBLP:conf/icml/YangLLZZW18} -- 
 research area. There are a variety of challenges, such as non-stationarity,  partial environment observability, and non-bounded system size.
In a case study, we have empirically showed the effectiveness of RL by showing a speed up of the convergence time of
 aggregate program we took into consideration, namely the hop count.
Furthermore, we have shown that the learnt program can easily scale up to system size.

To the best of our knowledge, this is the first effort towards a ``hybrid'' collective programming approach, where the high-level program specification is 
 done by a designer and the low-level implementation to building blocks based on machine learning.
%
As such, this work has shed light on several issues deserving future investigation.

Aggregate systems are partially observable from the perspective of each node. 
 Local observations may not be statistically significant in summarising the state of the entire aggregate. 
 In the future, we could address this problem through recurrent neural networks and a CTDE learning methodology.
 Learning algorithms were implemented at building block level. 
 Another possible direction is to use them at the framework level, e.g., to select which messages to send to the neighbourhood or not. 
 This has already been addressed in the literature under the name of \textit{learn to communicate}~\cite{DBLP:journals/corr/abs-1908-03963}. 
A novel way to handle this problem is to use Graph Neural Networks~\cite{DBLP:journals/corr/abs-1812-08434} to learn the representation of a distributed system and to select messages based on that representation. 
 Thus, we should consider this methodology when we will extend RL at the AC framework level.

In our experiments, we have applied a variation of the independent learning approach. 
 Agents first learn the best policy locally, and then at the end of the episode, a learner merges the Q-tables. 
 However, this does not reduce the problems of scalability and non-stationary, since a central agent must still collect all the Q-tables of the aggregate. 
 In particular, to handle the scalability issues, we will have to consider decentralised approaches based only on neighbour data (or local leaders).

Finally, we do not address the multi-agent credit assignment problem in this work. 
 Indeed, we use a global reward function for the whole aggregate. In this way, however, learning often degenerated because it was difficult to understand the influence of an agent on the system. 
 To reduce this problem, we plan to study the application of difference rewards~\cite{DBLP:conf/atal/AgoginoT04} or counterfactual baselines~\cite{DBLP:journals/corr/FoersterFANW17}.

\bibliographystyle{IEEEtran}
\bibliography{biblio-reduced}
%%
%% If your work has an appendix, this is the place to put it.
%\appendix

\end{document}

%%
%% End of file